### 2021-01-24 paper/intro.tex
First draft of Sec 1.2

% \begin{figure*}
%   \input{fig-sp}
%   \caption{Weakest Precondition and Strongest Postcondition \cite{}}
% \end{figure*}

Our approach follows that of weakest precondition semantics of
\citet{DBLP:journals/cacm/Dijkstra75}, which provides an alternative
characterization of Hoare logic \citep{Hoare:1969:ABC:363235.363259} by
mapping postconditions to preconditions.

% Our transformers $\aTr$ map preconditions to postconditions, in order to
% apply to preconditions.
% \begin{align*}
%   \hoare{\aForm}{\bCmd}{\bForm}
%   \;\;\Leftrightarrow\;\;
%   \hoare{\aTr(\aCmd,\aForm)}{\aCmd\SEMI\bCmd}{\bForm}
% \end{align*}
% For example
% \begin{gather*}
%   \hoare{r{=}1}{\PW{y}{r}}{y{=}1}
%   \\
%   \hoare{x{=}1}{\PR{x}{r}\SEMI\PW{y}{r}}{y{=}1}
% \end{gather*}

\endinput

This paper builds on
\cite{2019-sp} and
\cite{10.1145/3428262}.

Pomsets are a nice model of concurrent computation.
Predicate transformers are a nice model of sequential computation.
We show how these interact.

We do this by attaching predicate transformers to \emph{configurations}.

Thus the amount of transformation depends on which read events are seen in
the configuration.

We develop the theory of pomsets with predicate transformers, then apply the
theory to model relaxed memory.

Other things we could model include Map/Reduce and Fork/Join.

>>>>

\begin{figure*}
  \input{fig-sp}
  \caption{Weakest precondition semantics}
  \label{fig:sp}
\end{figure*}

Our approach follows that of weakest precondition semantics of
\citet{DBLP:journals/cacm/Dijkstra75}, which provides an alternative
characterization of Hoare logic \citep{Hoare:1969:ABC:363235.363259} by
mapping postconditions to preconditions. We recall the definition of
$\fwp{\aCmd}{\bForm}$ in Figure~\ref{fig:sp}. The rule we are most interested
in is the one for sequential composition, which maps seqential composition of programs
to function composition of predicate transformers.

Predicate transformers are a good fit to logical models of dependency calculation,
since both are concerned with preconditions, and how they are transformed by
sequential composition. Our first attempt is to associate a predicate transformer
with each pomset. We visualize this in diagrams by showing how $\bForm$ is transformed,
for example:
  \begin{align*}
    \begin{gathered}
      r \GETS x
      \\
      \hbox{\begin{tikzinline}[node distance=1ex and 1.5em]
          \event{rx0}{\DR{x}{0}}{}
          \xform{rx0d}{(r{=}0) \limplies \bForm}{below=of rx0}
          \xo{rx0}{rx0d}
        \end{tikzinline}}
    \end{gathered}
    &&
    \begin{gathered}
      s \GETS y
      \\
      \hbox{\begin{tikzinline}[node distance=1ex and 1.5em]
          \event{ry0}{\DR{y}{0}}{}
          \xform{ry0d}{(s{=}0) \limplies \bForm}{below=of ry0}
          \xo{ry0}{ry0d}
        \end{tikzinline}}
    \end{gathered}
    &&
    \begin{gathered}
      \IF{s{<}1} \THEN z\GETS r\&s \FI
      \\
      \hbox{\begin{tikzinline}[node distance=1ex and1.5em]
          \event{wz0}{(s {<} 1) \land (r\&s){=}0 \mid \DW{z}{0}}{}
          \xform{wz0d}{\bForm}{below=of wz0}
          \xo{wz0}{wz0d}
      \end{tikzinline}}
    \end{gathered}
  \end{align*}
For the sequentially consistent semantics, sequential composition composition
is straightforward: we apply each predicate transformer to the preconditions
of subsequent events, and compose the predicate transformers:
  \begin{gather*}
    r\GETS x\SEMI s\GETS y\SEMI \IF{s{<}1} \THEN z\GETS r\&s \FI
    \\
    \nonumber
    \hbox{\begin{tikzinline}[node distance=1ex and 1.5em]
        \event{rx0}{\DR{x}{0}}{}
        \event{ry0}{\DR{y}{0}}{right=of rx0}
        \event{wz0}{(r{=}0) \limplies (s{=}0) \limplies (s {<} 1) \land (r\&s){=}0 \mid \DW{z}{0}}{right=of ry0}
        \xform{rx0ry0d}{(r{=}0) \limplies(s{=}0) \limplies \bForm}{below=of ry0}
        \po{rx0}{ry0}
        \po{ry0}{wz0}
        \xo{rx0}{rx0ry0d}
        \xo{ry0}{rx0ry0d}
        \xo{wz0}{rx0ry0d}
      \end{tikzinline}}
  \end{gather*}
This model works for the sequentially consistent case, but needs to be
weakened for the relaxed case. The key observation of this paper is
that rather than working with one predicate transformer, we should
work with a \emph{family} of predicate transformers, indexed by sets
of events.

For example, for single-event pomsets, there are two predicate
transformers, since there are two subsets of any one-element set.
We call the predicate transfomer for $\emptyset$ the \emph{independent}
transformer, and the one indexed by $\{e\}$ the \emph{dependent} transformer.
We visualize this by including more than one transformed predicate,
with an edge leading to the dependent one. For example:
  \begin{align*}
    \begin{gathered}
      r \GETS x
      \\
      \hbox{\begin{tikzinline}[node distance=1ex and 1.5em]
          \event{rx0}{\DR{x}{0}}{}
          \xform{rx0i}{\bForm}{above=of rx0}
          \xform{rx0d}{(r{=}0) \limplies \bForm}{below=of rx0}
          \xo{rx0}{rx0d}
        \end{tikzinline}}
    \end{gathered}
    &&
    \begin{gathered}
      s \GETS y
      \\
      \hbox{\begin{tikzinline}[node distance=1ex and 1.5em]
          \event{ry0}{\DR{y}{0}}{}
          \xform{ry0i}{\bForm}{above=of ry0}
          \xform{ry0d}{(s{=}0) \limplies \bForm}{below=of ry0}
          \xo{ry0}{ry0d}
        \end{tikzinline}}
    \end{gathered}
    &&
    \begin{gathered}
      \IF{s{<}1} \THEN z\GETS r\&s \FI
      \\
      \hbox{\begin{tikzinline}[node distance=1ex and1.5em]
          \event{wz0}{(s {<} 1) \land (r\&s){=}0 \mid \DW{z}{0}}{}
          \xform{wz0i}{\bForm}{above=of wz0}
          \xform{wz0d}{\bForm}{below=of wz0}
          \xo{wz0}{wz0d}
      \end{tikzinline}}
    \end{gathered}
  \end{align*}
The model of sequential composition then picks which
predicate transformer to apply to an event's precondition by picking
the one indexed by all the events before it in causal order.

For example, we can recover the expected semantics
for the above example by choosing
the predicate transformer which is independent of $(\DR x0)$
but dependent on $(\DR y0)$, which is the transformer
which maps $\bForm$ to $(s{=}0) \limplies \bForm$.
  \begin{gather*}
    r\GETS x\SEMI s\GETS y\SEMI \IF{s{<}1} \THEN z\GETS r\&s \FI
    \\
    \nonumber
    \hbox{\begin{tikzinline}[node distance=1ex and 1.5em]
        \event{rx0}{\DR{x}{0}}{}
        \event{ry0}{\DR{y}{0}}{right=of rx0}
        \event{wz0}{(s{=}0) \limplies (s {<} 1) \land (r\&s){=}0 \mid \DW{z}{0}}{right=of ry0}
        \xform{rx0ry0d}{(r{=}0) \limplies(s{=}0) \limplies \bForm}{below=of ry0}
        \xform{rx0d}{(r{=}0) \limplies \bForm}{above=of rx0}
        \xform{ry0d}{(s{=}0) \limplies \bForm}{right=of rx0d}
        \xform{wz0i}{\bForm}{above=of wz0}
        \po{ry0}{wz0}
        \xo{rx0}{rx0d}
        \xo{ry0}{ry0d}
        \xo{rx0}{rx0ry0d}
        \xo{ry0}{rx0ry0d}
      \end{tikzinline}}
  \end{gather*}
